Finding communities in graphs

### 2012-12-15 CMPT771/project_report/paper.tex
almost done

%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\frenchspacing
\pdfinfo{
/Title (Formatting Instructions for Authors Using LaTeX)
/Subject (AAAI Publications)
/Author (AAAI Press)}
\setcounter{secnumdepth}{1}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Probability Based Partitioning of Online Social Networks}
\author{Vivek Mahajan\\
vmahajan@sfu.ca\\
}
\maketitle
\begin{abstract}
\begin{quote}
Online social networking has  become a phenomenon in the last few years. The
growing usage of social networks has put the service providers under a lot of
pressure to efficiently process the users queries.  In social networks like
orkut, most of the queries access the immediate neighbors of the current node
and neighbors of its neighbors.  Existing graph partitioning approaches focus
on reducing the number of inter server links, neglecting the query patterns.
Our algorithm focuses on partitioning the graph so as to minimize the number of
servers to be accessed in order to process a query. Our results show that the
proposed algorithm gives useful partitioning of social networks, thereby
reducing inter server accesses.

@@

\section{Introduction}
    In the last few years, the online social networks(OSNs) have grown in huge
proportion with respect to the number of users as well as its utility. The
amount of usage of OSNs exceeds e-mail usage nowadays, which is remarkable.
OSNs are the most popular sites on the Internet. A
study\footnote{http://www.news.com.au/technology/story/0,28348,25170441-5014239,00.html}
reported that OSNs account for 10\% of all the Internet time worldwide. The
growth can be appreciated by the fact that there was 1382\% increase in the
user base of twitter in one month[Feb-Mar 09]
\footnote{http://blog.nielsen.com/nielsenwire/onlinemobile/twitters-tweet-smell-of-success}.
There can be variations of OSNs in terms of usability e.g On orkut, the users
interact with each other by exchanging scraps or photographs, while on
SLASHDOT, the users interact via blogs. Facebook, LinkedIn are other examples
of OSNs.

    The phenomenal growth in the size of OSNs has given rise to scalability
issues that can be solved by distributing online social network graph among
different servers. However the basis of partitioning  is a major issue. That
is, the graph partitioning could be based on structural, geo-political or
traffic properties, which are discussed in detail in \cite{arxviv} for twitter
and orkut. In this paper, we introduce a new partitioning criterion called
\texttt{hop based partitioning} which takes into consideration the typical
nature of social network queries.

\subsection{Hop Based Partitioning}

Social network queries typically visit their one hop and two hop neighbors. The
small world phenomenon links any two users on a social network with a distance
of six hops. Hence, while partitioning a social network, it would be beneficial
to maintain the one-hop and two-hop neighbors of a frequented query node on the
same server\footnote{through the rest of the paper the server and partition
terms are used interchangeably} as itself. Further, it is expected that any
graph partitioning algorithm attempts to reduce the inter server edges and
equally distribute the data across servers.

@@

    In this project, we propose a novel probability based objective function
and  an algorithm to find the clusters which optimizes that. Given a social
network graph $G$, our algorithm partitions the graph while meeting the
following constraints:
\begin{itemize}
\item For any node in a partition most of its two-hop neighbors lie within the same partition.
\item All partitions are approximately equally sized.
\end{itemize}

@@

    In this paper, we  compare our results with METIS\cite{20}. METIS focuses
on minimizing the cuts between the partitions as well as into roughly equal
partitions. The METIS algorithm is divided into three phases termed as
coarsening, partitioning and uncoarsening/refinement.  The main motivation
behind the algorithm is to generate an approximate graph of the original graph,
partitioning it and then retracting the coarsening phase while doing the
refinement between each iteration.  Our algorithm also follows a hierarchical
clustering approach where repeated merging of clusters is done for some
iteration and then the graph randomly partitioned with imposed contraints of
having nearly equal partitions. Finally, the refinement algorithm is applied on
the clusters thus obtained. We differ from METIS, in terms of our partitioning
criterion, our novel way of clustering and partitioning, and  the modification
and the method of refinement applied which is applied only once.

\section{Probability Based Objective Function}

   To calculate our objective function, we assume that there is an equal chance
of each user of an OSN to come online. However, we can easily modify our
objective function if we know the probability of each user coming online.

 Given some partitions, an evaluation criteria should be able to give a good
index of the quality of partitions in terms of fraction of one and two hops
lying outside the partition and should also give importance to the past
interaction between the users of OSNs.

This index calculates the probability of one and two hops from a node to lie in different partition.

Formally, let $X$ = $\{x_1,x_2,x_3,\cdot\cdot\cdot,x_n\}$ be the nodes in an OSN, $x_i$ = $i^{th}\ node$, 
$\{y_1, y_2, y_3,\cdot\cdot\cdot,y_j\}$ be the $j$ neighbors of $x_i$. Let $\{v_1, v_2, v_3, \cdot\cdot\cdot,
 v_j\}$ be the number of interactions between $x_i$ and its neighbors. And $V_i = \sum_{k\in j} v_k$.
 Let $Y^\prime$ be 
the subset of neighbors of $x_i$ which lie in partition other than $x_i$'s. Let $Y^\prime = \{y_1^\prime, y_2^\prime, \cdot\cdot, y_k^\prime\}$. Let the interaction between $x_i$ and the users in this subset be  $V^\prime = \{v_1^\prime, v_2^\prime, \cdot\cdot, v_k^\prime\}$.
$V_i^\prime = \sum_{p\in k} v_p^\prime$.

        Let $P_{x_i}^{1-hop}$ be the probability of one-hop neighbor of $x_i$ to lie in separate partition, and $P_{x_i}^{2-hop}$ be probability of two hops from $x_i^{th}$ node to lie in the separate partition.\\
        Clearly, $P_{x_i}^{1-hop}= \frac{V_i^\prime}{V_i}$ \\
        $P_{x_i}^{2-hop}=\frac{1}{Number\ of\ neighbors\ of\ x_i} * \displaystyle\sum_{k=1}^j P_{x_i y_k}$, \\
        where,\\
        $P_{x_i y_k}$ = $\frac{v_k}{V_i}$* $P_{x_{k}}^{1-hop}$ =  Probability of two hop from  $x_{i}$ to lie  in a  partition other than  its own through $y_k$, if $x_i$ and $y_{k}$ are in the same partition.\\
        $P_{x_i y_k}$ =  1 , if $x_i$ and $y_k$ are in the separate partition, as we assume if the one hop is on a different partition, it is almost
certain that two hop lie in the same partition.\\
        $Probability\ Index=\frac{1}{Total\ number\ of\ nodes}*\displaystyle\sum_{i=1}^nP_{x_i}^{2-hop}$, $n$ is the total number of nodes.\\

        This probability index gives us an overall probability of a two hop
query from a node to lie in the different partition. In the following section
we will propose an algorithm which optimizes this objective function.

@@

In this section, we introduce our graph partitioning algorithm. A social
network  partitioning algorithm is expected to reduce inter server accesses
requiring neighbors and the neighbors of neighbors of a query nodes to be
present in the same server as itself.  To reduce the partitioning overhead, it
is important that inter server links be reduced while trying to maximize the
number of nodes for which its one and two hop neighbors be on the same server
as itself.  Further, in order to equally distribute the workload across servers
and optimally use the servers used, it is essential that the partitions size
should be approximately equal. Our algorithm attempts to fulfil the above
requirements while attempting to maximize the number of nodes for which a
majority of its one-hop neighbors and with next priority its 2-hop neighbors
are included within the same server.  Our algorithm is divided into three
phases, namely the graph coarsening phase, partitioning phase and the
refinement phase. The graph coarsening phase reduces the size of the original
graph by some factor by clustering the nodes \emph{intelligently}.  A weight
function is used in order to decide which nodes should be clustered  during the
coarsening phase. In the partitioning phase, we randomly generate partitions of
the coarsened graph. The refinement phase next reduces the inter partition
edges by making slight modifications to the partitioned graph. We next give
definitions of commonly used terms followed by a detailed description of each
of the phases.

@@

    We modify one of the phases of  Kernighan-Lin algorithm for our criterion.
To be precise, we change the definition of $\delta$$Q$ for us. In the
Kernighan-Lin algorithm, $\delta$$Q$ is the change in the edges cuts between
the partitions when the nodes are swapped. Let $a$, $b$ be two nodes that lie
in different partitions. Let $I_a$ be the internal cost for $a$, i.e the
neigbors of $a$ that lies in the same partition as of $a$ and  $E_a$ be the
external cost for the node $a$, i.e the number of neighbors of $a$ that lie in
the separate partition. Similarily $E_b$ and $I_b$ are the external cost and
internal cost for the node $b$. Now let $D_a$ = $E_a$ - $I_a$ and $D_b$ = $E_b$
- $I_b$ , i.e the difference of external and internal costs. If $a$ and $b$ are
interchanged then the gain i.e $\delta$$Q$ will be $D_a + D_b -2C_{a,b}$. We
make changes in the internal, external and $C_{a,b}$ to fit our purpose of
decreasing the probability of the two hops from one node lying in the different
partitions. We define internal cost $I_{ap}$ as the probability of an edge of
node $E_{ap}$ lying in the same partition as of $a$ ,external cost as
probability of an edge of node $a$ lying in a separate partition. Hence,\\
\begin{displaymath}I_{a}=\frac{I_a}{Degree\ of\ node\ a }\\\end{displaymath},
\\\begin{displaymath}E_{a}=\frac{E_a}{Degree\ of\ node\ a}\\\end{displaymath}
and\\ \begin{displaymath}2C_{a,b}=\frac{1}{Degree\ of\ node\
a}+\frac{1}{Degree\ of\ node\ b}\\\end{displaymath}

@@

\section{Experiments and Results}

\section{Conclusion}
\bibliographystyle{aaai} \bibliography{xbib}
\end{document}

>>>>

%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\frenchspacing
\pdfinfo{
/Title (Formatting Instructions for Authors Using LaTeX)
/Subject (AAAI Publications)
/Author (AAAI Press)}
\setcounter{secnumdepth}{1}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Hop Probability Based Partitioning of Online Social Networks}
\author{Vivek Mahajan\\
vmahajan@sfu.ca\\
}
\maketitle
\begin{abstract}
\begin{quote}
Online social networking has  become a phenomenon in the last few years. The
growing usage of social networks has put the service providers under a lot of
pressure to efficiently process the users queries.  In social networks like
facebook, most of the queries access the immediate neighbors of the current node
and neighbors of its neighbors.  Existing graph partitioning approaches focus
on reducing the number of inter server links, neglecting the query patterns.
Our algorithm focuses on partitioning the graph so as to minimize the number of
servers to be accessed in order to process a query. Our results show that the
proposed algorithm gives useful partitioning of social networks, thereby
reducing inter server accesses.

@@

\section{Introduction}
    In the last few years, the online social networks(OSNs) have grown in huge
proportion with respect to the number of users as well as its utility. The
amount of usage of OSNs exceeds e-mail usage nowadays, which is remarkable.
OSNs are the most popular sites on the Internet. A
study\footnote{http://www.news.com.au/technology/story/0,28348,25170441-5014239,00.html}
reported that OSNs account for 10\% of all the Internet time worldwide. The
growth can be appreciated by the fact that there was 1382\% increase in the
user base of twitter in one month[Feb-Mar 09]
\footnote{http://blog.nielsen.com/nielsenwire/onlinemobile/twitters-tweet-smell-of-success}.
There can be variations of OSNs in terms of usability e.g On facebook, the users
interact with each other by posting on the walls or sharing photographs, while on
SLASHDOT, the users interact via blogs. LinkedIn and Google+ are other examples
of OSNs.

    The phenomenal growth in the size of OSNs has given rise to scalability
issues that can be solved by distributing online social network graph among
different servers. However finding the basis of partitioning  is a major issue. That
is, the graph partitioning could be based on structural, geo-political or
traffic properties, which are discussed in detail in \cite{arxviv} for twitter
and orkut. In this paper, we introduce a new partitioning criterion called
\texttt{Hop probability based partitioning} which takes into consideration the typical
nature of social network queries.

\subsection{Hop probability Based Partitioning}

Social network queries typically visit their one hop and two hop neighbors. The
small world phenomenon links any two users on a social network within a distance
of six hops. Hence, while partitioning a social network, it would be beneficial
to maintain the one-hop and two-hop neighbors of a frequented query node on the
same server\footnote{through the rest of the paper the server and partition
terms are used interchangeably} as itself.

@@

    In this project, we propose a novel hop probability based objective function
and  an algorithm to find the clusters which optimizes that. Given a social
network graph $G$, our algorithm partitions the graph while meeting the
following constraints:
\begin{itemize}
\item For any node in a partition most of its two-hop neighbors lie within the same partition.
\item All partitions are approximately equally sized.
\end{itemize}

@@

    In this paper, we  compare our results with METIS\cite{20}. METIS focuses
on minimizing the cuts between the partitions as well as into roughly equal sized
partitions. The METIS algorithm is divided into three phases termed as
coarsening, partitioning and uncoarsening/refinement.  The main motivation
behind the algorithm is to generate an approximate graph of the original graph,
partitioning it and then retracting the coarsening phase while doing the
refinement between each iteration.  Our algorithm also follows a hierarchical
clustering approach where repeated merging of clusters is done for some
iteration and then the graph is randomly partitioned with imposed contraints of
having nearly equal partitions. Finally, the refinement algorithm is applied on
the clusters thus obtained. We differ from METIS, in terms of our partitioning
criterion, our novel way of clustering, and  the modified form Kerinighan-Lin 
algorithm which  refines the generated partitions.

\section{Hop probability Based Objective Function}

   To calculate our objective function, we assume that there is an equal chance
of each user of an OSN to come online. However, we can easily modify our
objective function if we know the probability of each user to come online.

 Given partitions of a graph, our evaluation criteria gives a good
quantitative measure of the quality of partitions in terms of fraction of one and two hops
lying outside the partition and should also give importance to the past
interaction between the users of OSNs. Essentially, our objective function
calculates the average probability of one and two hops from a node to lie in different
partition.

Formally, let $X$ = $\{x_1,x_2,x_3,\cdot\cdot\cdot,x_n\}$ be the nodes in an OSN, $x_i$ = $i^{th}\ node$, 
$Y_i=\{y_1, y_2, y_3,\cdot\cdot\cdot,y_j\}$ be the $j$ neighbors of $x_i$. Let $V_i=\{v_1, v_2, v_3, \cdot\cdot\cdot,
 v_j\}$ be the number of interactions between $x_i$ and its neighbors. And $|V_i| = \sum_{k\in j} v_k$.
 Let $Y^\prime_i$ be 
the subset of $Y_i$ which lie in partition other than $x_i$'s. Let $Y^\prime_i = \{y_1^\prime, y_2^\prime, \cdot\cdot, y_k^\prime\}$. 
Let the interaction between $x_i$ and the users in this subset be  $V^\prime_i = \{v_1^\prime, v_2^\prime, \cdot\cdot, v_k^\prime\}$.
$|V_i^\prime| = \sum_{p\in k} v_p^\prime$.

        Let $P_{x_i}^{1-hop}$ be the probability of one-hop neighbor of $x_i$ to lie in separate partition, and $P_{x_i}^{2-hop}$ be probability of two hops from $x_i^{th}$ node to lie in the separate partition.\\
        Clearly, $P_{x_i}^{1-hop}= \frac{|V_i^\prime|}{|V_i|}$ \\
        $P_{x_i}^{2-hop}=\frac{1}{Number\ of\ neighbors\ of\ x_i} * \displaystyle\sum_{k=1}^j P_{x_i y_k}$, \\
        where,\\
        $P_{x_i y_k}$ = $\frac{v_k}{|V_i|}$* $P_{x_{k}}^{1-hop}$ =  Probability of two hop from  $x_{i}$ to lie  in a  partition other than  its own through $y_k$, if $x_i$ and $y_{k}$ are in the same partition.\\
        $P_{x_i y_k}$ =  1 , if $x_i$ and $y_k$ are in the separate partition, as we assume if the one hop is on a different partition, it is almost
certain that two hop lie in the same partition.\\
        $Probability\ Index=\frac{1}{Total\ number\ of\ nodes}*\displaystyle\sum_{i=1}^nP_{x_i}^{2-hop}$, $n$ is the total number of nodes.\\

        This probability index gives us an overall probability of a two hop
query from a node to lie in the different partition. To the best of our
knowledge this kind of evaluation of evaluation criteria has never been
published before. This is quite novel in the sense it takes into account the
user access patterns.  In the following section we will propose an algorithm
which optimizes this objective function.

@@

In this section, we introduce our graph partitioning algorithm. A social
network  partitioning algorithm is expected to reduce inter server accesses,
requiring friends and the friends of friends of a query node to be present in
the same server as itself.   Also, in order to equally distribute the workload
across servers , it is essential that the partitions size should be
approximately equal.

 Our algorithm attempts to fulfil the above requirements
while attempting to maximize the number of nodes for which a majority of its
one-hop neighbors and with next priority its 2-hop neighbors are included
within the same server.  Our algorithm is divided into three
phases, namely the graph coarsening phase, partitioning phase and the
refinement phase. The graph coarsening phase reduces the size of the original
graph by some factor by clustering the nodes \emph{intelligently}.  A weight
function is used in order to decide which nodes should be clustered  during the
coarsening phase. In the partitioning phase, we randomly generate partitions of
the coarsened graph. Due to time constraint we decided to randomly generate the
partitions. Alternatively, we can use traditional graph partitioning methods.
In the next phase, which we call refinement phase we reduce the inter partition
edges by using a Kernighan-Lin algorithm. We next give definitions of commonly
used terms followed by a detailed description of each of the phases.

@@

    We modify one of the phases of  Kernighan-Lin algorithm for our criterion.
To be precise, we change the definition of $\delta$$Q$ for us. In the
Kernighan-Lin algorithm, $\delta$$Q$ is the change in the edges cuts between
the partitions when the nodes are swapped. Let $a$, $b$ be two nodes that lie
in different partitions. Let $I_a$ be the internal cost for $a$, i.e the total
number of interactions between $a$ and its friends which lie in the same
partition and  $E_a$ be the external cost for the node $a$, i.e the total
number of interaction between $a$ and its friends which lie in the separate
partition. Similarily $E_b$ and $I_b$ are the external cost and internal cost
for the node $b$. Now let $D_a$ = $E_a$ - $I_a$ and $D_b$ = $E_b$
- $I_b$ , i.e the difference of external and internal costs. If $a$ and $b$ are
  interchanged then the gain i.e $\delta$$Q$ will be $D_a + D_b -2C_{a,b}$. We
make changes in the internal, external and $C_{a,b}$ to fit our purpose of
decreasing the probability of the two hops from one node lying in the different
partitions. We define internal cost $I_{ap}$ as the probability of an edge of
node $E_{ap}$ lying in the same partition as of $a$ ,external cost as
probability of an edge of node $a$ lying in a separate partition. Let $T_a$ be
$Total\ number\ of\ interactions\ done\ by\ user\ a$ and $T_b$ be $Total\
number\ of\ interactions\ done\ by\ user\ b$. And $I_{ab}$ be the number of
interactions between $a$ and $b$. Hence,\\
\begin{displaymath}I_{ap}=\frac{I_a}{T_a}\\\end{displaymath},
\\\begin{displaymath}E_{ap}=\frac{E_a}{T_a}\\\end{displaymath} and\\
\begin{displaymath}2C_{ap,bp}=\frac{I_{ab}}{T_a}+\frac{I_{ab}}{T_b}\\\end{displaymath}

@@

\begin{table*}[htb]
\centering
\caption{Datasets - statistics}
\begin{tabular}{|c|c|c|c|}
\hline
 Datasets& Nodes & Edges \\
\hline
Facebook& 63731 & 68885  \\
\hline
SOC-Epinions & 75879 & 508837   \\
\hline
SLASHDOT& 77360 & 905468  \\
\hline
\end{tabular}

\label{fig:twopartcuts}
\end{table*}
\begin{table*}[htb]
\centering
\caption{Two Partitions - Evaluation Indexes}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
 &\multicolumn{2}{c|}{Our Algorithm}&\multicolumn{2}{c|}{METIS}\\
\hline
%\cline{2-5}
 DataSets&Hop Probability score &Edge Cuts&Hop Probability Score&Edge Cuts\\
\hline
Facebook& 0.225316 & 0.273173  & 0.21507 & 0.269943\\
\hline
SOC-Epinions & 0.135977 & 0.304285  & 0.162591 & 0.313931\\
\hline
SLASHDOT& 0.37827 & 0.496867  & 0.438342 & 0.762978\\
\hline
\end{tabular}
\label{fig:twopartindex}
\end{table*}

\section{Experiments and Results}
We have picked three popular and reasonaably sized datasets to run our experiments on
\begin{itemize}
\item \textbf{Facebook Social Network}: Facebook is a true social networking
site on which users interact by writing on the wallposts and sharing lots of
media content. The dataset was collected between December 29th, 2008 and
January 3rd, 2009, and between January 20th 2009 and January 22th 2009 from the
Facebook New Orleans network. The original dataset is time stamped, mentioning
when were the links made.  For our purposes we used it as an undirected graph
giving no consideration to time stamp attribute.  We also have list of all the
wall posts for Facebook New Orleans network. The dataset was anonymized and we
only have information about which user posted on whose wall at what time. No
posts content is provided. But this information was good enough to find the
interactions between the users which forms the weights of the edges between the
users.  So we build a graph in which nodes are represented by users and there's
and edge between two nodes if those two users are friends and weights are the
average number of interactions between the two users i.e average number of
times they posted on each other's wall + 1. 1 is added as there were cases when
two users were friends but they didn't interact with each other. So we decided
on having edge weight as 1 by default. As atleast they interacted once when
they accepted the friends request. The dataset was crawled by \cite{26}.

\item \textbf{Epinions Social Network}: This is who-trust-whom online social
network of a general consumer review site Epinions.com. Members of the site can
decide whether to "trust" each other.  All the trust relationships interact and
for the web of trust which is then combined with review rating to determine
which reviews are shown to the user. We construct a graph in which the users
are represented by nodes and there is an edge between a user if they "trust"
each other. Since we didn't have any user query access pattern information for
this graph, we gave each edge a weight of 1. The
dataset\footnote{http://snap.stanford.edu/data/soc-Epinions1.html} was first
used in \cite{epi2003}.

\item \textbf{Slashdot}: Slashdot is a technology related news website known
for its specific user community.  The website features user submitted and
editor evaluated current primarily technology oriented news.  In 2002 Slashdot
introduced the Slashdot Zoo feature which allows users to tag each other as
neighbors or foes.  The network contains friend/foe links between the users of
Slashdot. Again, the users are nodes and a link between them indicates whether
they are friend or foe. And edge weight is set to 1 as we don't have any access
pattern information. The
dataset\footnote{http://snap.stanford.edu/data/soc-Slashdot0811.html} was first
introduced in \cite{25}.

\end{itemize}

We compare the hop probability based partitioning approach with METIS. We are
considering only bipartitions in our results. The input graph to the two
algorithms are same except the facebook graph. For our algorithm, the input
Facebook graph is with edge weights based on  interactions between users, while
input Facebook graph for METIS is with all the edge weights set to 1. The main
reason is that METIS doesn't take edge weights as input, so we can't provide it
with that. We also have a reasonable contraint that the partitions should be of
roughly equal size. We compare the two algorithms on edges cuts and hop
probability based evaluation criteria. As we can see the periliminary results
are quite close to METIS in terms of edge cuts as well the novel criteria. We
think that some more experimentation with the weight function, partitioning
algorithm(here we are doing random partitioning) and the refinement algorithm,
might result in better results. 
\section{Conclusion and Future Work}
In this paper, our focus was to partition an Online Social Network graph such
that the probability of query(one-hop/two-hop) generated from a user lying in
different partition is reduced. We proposed a novel criteria for evaluating the
partitions and presented a hierarchical algorithm optimizing that criteria. Our
preliminary results looks promising. We plan to further improve it by 
analyzing the graph formation at each level and also by experimenting with the
weight function as well as the partitioning function. Refinement algorithm can
be changed as well. As we don't know whether the modified Kernighan Lin
algorithm will surely converge. I plan to work on this a little more and maybe
submit it as a short paper in some conference/workshop.
\bibliographystyle{aaai} \bibliography{xbib}
\end{document}

### 2012-12-15 CMPT771/survey_report/paper.tex
almost done

\section{Introduction}

\section{Objective Functions}
If you have clusters of a given graph as an output of a clustering algorithm, we need to have some quantitative measure of how "good" they are. We call them objective
function which the algorithms try to optimize. The intution behind most of the objective functions is that a good cluster has many edges internally and few edges 
pointing outside. In this section, we look at the other objective functions that are based on this intution.

@@

We consider the following metrics $f(S)$ that capture the notion of a quality of the cluster. Lower value of score $f(S)$ signifies a more community-like set of 
nodes.
\begin{itemize}
\item \textbf{conductance}: $f(S) = \frac{c_S}{c m_S+c_S}$ measures the fraction of edges that points outside the
cluster
\item \textbf{Expansion}: $f(S) = \frac{c_S}{n_S}$ measures the number of edges per node that point outside the
cluster
\item \textbf{Internal density}: $f(S) = 1 - \frac{m_S}{n_S(n-n_S)}$ is the fraction of all possible edges leaving
the cluster
\item \textbf{Cut Ratio}: f(S) = $\frac{c_S}{n_S(n-n_S)}$ is the fraction of all possible edges leaving the cluster 
\item \textbf{Normalized Cut}:  f(S) = $\frac{c_S}{2m_S+c_S} + \frac{c_S}{2(m-m_S)+c_S}$
\end{itemize}

>>>>

\section{Introduction}
The origin of graph theory dates back to Euler's solution of the puzzle of
Konigsberg's bridges in 1736.  Since then a lot has been learned about graphs
and their mathematical properties. It is arguably the most widely adopted
representation which can be used to abstract lots of  real world problems. The
problems in Biology, sociology and information networks can be formulated as
graphs, and thus the graph analysis has become crucial to understand the
features of these systems. Recently, with faster processors and plenty of
computational resources, researchers can process and analyze huge amount of
data, which wasn't possible in earlier times.

Graphs representing real systems are more often not regular graphs. Some are
random graph with higly homogeneous degree distribution. Real networks are not
random graphs, as they display inhomogeneities, revealing a high level of order
and organization, in which high degree vertices can co-exist with low degree
vertices. Furthermore, the distribution of edges is not only globally, but also
locally inhomogeneous, with high number of edges within some group of vertices,
and low number of edges between these groups. This special feature of real
networks is called community structure or clustering. Communities, also called
clusters or modules, are groups of vertices which most likely share some common
properties.

\begin{figure}
%\begin{minipage}[htb]{0.48\linewidth}
\includegraphics[scale=0.40]{pics/intro.eps}
\caption{A simple graph with three communities, enclosed by the dashed circles.}
\label{fig:bipart}
%\end{minipage}
\end{figure}

The communities exists in society, like families, working and friendship
circles, towns, nations etc.  The ubiquity of Internet has also resulted into
creation of virtual groups like online communities in social networks like
Facebook or Linkedin. Communities also occur in many networked systems from
biology, computer science, economics, politics etc.  In protein-protein
interaction networks, communities are likely to group proteins having the same
functional property in the cell. In WWW the may correspond to groups of pages
dealing with same related topics, and so on.

In WWW, clustering users who hvae similar interests and who are geografically
near to each other may improve the performance of services provided on the
World Wide Web, where each cluster of clients could be served by a dedicated
mirror server. Clustering can be used  to generate reccomendations on
e-commerce websites, e.g users with similar interests can be clustered and
could be given appropriate reccomendation.  Clusters in large sparse graphs can
be used to create data structures in order to efficiently store the graph data
and to handle navigational queries, like path searches. It can also used in
case of Ad hoc networks, as they normally don't maintain a central routing
tables. Also, clustering can be used in case of caching of web pages as well.

The goal of detecting communities in graphs is to identify the clusters and,
more often, their hierarchical organization, by using the structure properties
of the graph. The problem itself is very old and have lots of proposed
algorithms in different disciplines.

The first analysis of community structure was carried out by Weiss et al. in
1957, who searched for group of workers in a government agency. Even earlier
work was done in finding cluters of people in small political bodeis, based on
the similarity of their voting pattern.

Finding graph communities is also popular in computer science. In parallel
computing, it is quite common to find out the best way to allocate tasks to
processors so as to minimize the communications between them. This is also
called graph partitioning problem which will be dicussed in the later sections.
In a seminal paper appeared in 2002, Girvan and Newman proposed a new algorithm
for finding the edges lying between communities and iteratively removing them
to find the clusters \cite{new2002}. After this paper, the field was revived and people
gave lots of sophisticated algorithms to tackle this problem.

In this survery, we try to cover some of the most popular algorithms in this
area. We start with giving some examples of communities in Real-world networks.
In section 2, we will give definition of various objective functions used
nowaday. We will continue the discussion with explaining one of the most widely
used objective function in section 3. Followed by explanation of some
traditional approaches in section 5, 6, 7 and 8. We will then give a brief
overview of a more recent method in section 9 and then we will conclude in
section 10.
\subsection{Communities in Real-World Networks}
\section{Objective Functions}
If you have clusters of a given graph as an output of a clustering algorithm,
we need to have some quantitative measure of how "good" they are. We call them
objective function which the algorithms try to optimize. The intution behind
most of the objective functions is that a good cluster has many edges
internally and few edges pointing outside. In this section, we look at the
other objective functions that are based on this intution.

@@

We consider the following metrics $f(S)$ that capture the notion of a quality of the cluster. Lower value of score $f(S)$ signifies a more community-like set of 
nodes.
\begin{itemize}
\item \textbf{conductance}: $f(S) = \frac{c_S}{c m_S+c_S}$ measures the fraction of edges that points outside the
cluster \cite{good04,shi2000}
\item \textbf{Expansion}: $f(S) = \frac{c_S}{n_S}$ measures the number of edges per node that point outside the
cluster \cite{radi04}
\item \textbf{Internal density}: $f(S) = 1 - \frac{m_S}{n_S(n-n_S)}$ is the fraction of all possible edges leaving
the cluster \cite{radi04}
\item \textbf{Cut Ratio}: f(S) = $\frac{c_S}{n_S(n-n_S)}$ is the fraction of all possible edges leaving the cluster 
\item \textbf{Normalized Cut}:  f(S) = $\frac{c_S}{2m_S+c_S} + \frac{c_S}{2(m-m_S)+c_S}$ \cite{shi2000}
\end{itemize}
